Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs

### 2014-06-26 Referee1_Response.tex

\begin{itemize}
\item[13.] \textit{p14, l35: The authors have gone to some length to distinguish different types of $A_j$ terms.
However, this equation has an $\mathbb{E}A_j^2$ term that has been collapsed from a sum, but we don't know ``which $j$''. 
Should there be a $\max_j \mathbb{E}A_j^2$ or some other treatment here?}
\end{itemize}

\noindent xxx NEED TO CHECK AND CORRECT AS NECESSARY. 
\medskip

\begin{itemize}
\item[14.] \textit{p15, Theorem 3: A hypothesis in the theorem is that $\hat{x} \neq x^*$, and yet the proof starts with the case that $\hat{x} = x^*$. 
That hypothesis could be removed if the claim had a ``$\geq$'' or this statement could be removed or modified.}
\end{itemize}

\noindent Thank you for pointing this out.  
We removed this statement from the proof of Theorem 3 and added it as a remark to the paragraph after the theorem.
\medskip

@@

\begin{itemize}
\item[16.] \textit{p17, l55: Perhaps add a comment on why DB1 may not experience the asymptotic variance reduction? 
(Is it because it may lack a unique solution?)}
\end{itemize}

\noindent
Done.
% xxx Yes, it is partially because we are not sure if it has a unique optimal solution as DB1 has never been solved before. 
%Even if it has been solved, the variance reduction is shown asymptotically but the computations are only for small sample sizes. 
%We reworded this to reflect these points.
%Thank you for this suggestion. 
\medskip

\begin{itemize}
\item[17.] \textit{p18, figures: The figures are a bit small and hence can be hard to read without enlarging them (but perhaps this will be taken care of at production).}
\end{itemize}

\noindent 
We split these two figures into three and enlarged each graph. 
We also enlarged the legends in these graphs to improve readability.
All the new graphs added to the paper in Section 5.4 (Computational Efficiency) are also larger, with legible legends. 
\medskip

@@

\begin{itemize}
\item[18.] \textit{p18, l57: The authors speak of coverage probabilities for DB1, but how can this be done without knowing an optimal solution to DB1?}
\end{itemize}

\noindent 
We used a Latin hypercube sample of size 50,000 estimate $\mu_{\hat{x}}$ for DB1, which was then used to estimate the coverage probability of the OMRP confidence intervals for this test problem.
We explain this at the end of Section 5.1 (Test Problems). 
We also changed the wording in Section 5.3 (Results of Experiments) to make it clear that the coverage probability is estimated for DB1. 
\medskip

@@

\begin{itemize}
\item[19.] \textit{p19: It may be possible to expand on the last paragraph of Section 5, in giving guidance on selecting the degree of overlap, depending on the authors' response to comment M1 above.}
\end{itemize}

\noindent 
Thank you for this suggestion. 
Based on our study of computational efficiency of the OMRP estimators, we  provided an updated and expanded guidance on selecting the degree of overlap. 
This guidance appears in the third and last paragraphs of the new section, \S 5.4 (Computational Efficiency).

### 2014-06-26 omrp_lb_rev.tex
Finer pass through Ref2 comments - PART II

We wish to assess the quality of a candidate solution $\xh \in X$. 
Assessing solution quality is important in practice because as (\ref{eq:sto_prog}) typically cannot be solved exactly, one only has an approximate solution $\xh$ without verification of its quality.
Assessing solution quality is also a critical component of stopping criteria in algorithms designed to (approximately) solve (\ref{eq:sto_prog}).
We define the quality of $\xh$ by its optimality gap, $\e{f(\xh,\xit)} - \zs$. 
The lower the optimality gap, the higher the quality of the solution; and a zero optimality gap implies that the solution is optimal. 
Note that the candidate solution $\xh \in X$ may have been found in any way. 
We assume that this solution is given as input to our procedures, and hence fixed, throughout this paper.

@@

Our aim is to apply the idea of overlapping batches to MRP.  
Overlapping could help to obtain better variance estimators with simple data reuse, resulting in improved CIs.
%Especially for problems where sampling error is dominant, this simple data reuse method could yield better estimators.
We note several differences in this setting compared to the classical simulation setting.  
In simulation output analysis, the point of interest is that of estimating the variance of the sample mean of a covariance stationary process.  
We are interested in estimating the variance of an optimality gap estimator.  
Notice that an optimality gap estimator not only has a sample mean $(\frac{1}{m} \sum_{i=1}^m f(\xh,\xiti))$ but also a minimized sample mean $(\zs_m)$.  
Minimization changes the statistical properties of sample means.  
For instance, the central limit theorem may not hold for a minimized sample mean, even though it holds for each $\x \in \X$.  
Therefore, analysis of the asymptotic benefits of overlapping becomes more complicated.
We overcome this technical difficulty by approximating the optimality gap estimators by their nonoptimized counterparts (see \S \ref{sec:theory}).  
The nonoptimized counterparts are (regular) overlapping batch means estimators (without any optimization present); therefore, they have the desired statistical properties of approximately the same bias and lower variances compared to their nonoverlapping versions. 
We establish convergence of the optimality gap estimators that use overlapping to their nonoptimized counterparts. 
Another difference between our setting and simulation output analysis is that once the data are generated through a simulation, they can be reused without much additional computational effort to obtain the OBM variance estimator.  
In this setting, due to the need to solve a sampling problem (SP$_m$), the computational effort may increase with data reuse.  
This can be alleviated in two ways. 
First, near-optimal variance reduction can be obtained by partially overlapping the batches \citep{Welch1987,Song1992}.  
Partial overlapping results in a smaller number of batches; hence, fewer optimization problems need to be solved.
In this paper, we consider variable amounts of overlap and examine this both theoretically and empirically. 
Second, in many solution methods, warm-starting can be used to quickly solve the sampling approximations with overlapping samples, considerably reducing solution time.

@@

Nonoverlapping batch means takes $n$ observations, $y^1, \dots, y^n$ of the stochastic process and splits them into $k$ batches of size $m$, where $k = \frac{n}{m}$ (for simplicity, assume, for now, $n = mk$).  
The sample mean of each batch is then computed as
$
	\yb_j = \frac{1}{m} \sum_{i=1}^{m} y^{m(j-1)+i}, j = 1,2, \dots, k,
$
and the overall sample mean $\ybb = \frac{1}{k} \sum_{j=1}^k \yb_j = \frac{1}{n} \sum_{i=1}^n y^i$ provides a point estimator of $\mu$.  
The variance estimator of the batch means, which is $n/m$ times the sample variance, is calculated as
\begin{equation} \label{eq:var}
	\varhat(\ybb) = \frac{m}{n}\frac{1}{k-1} \sum_{j=1}^k \left( \yb_j - \ybb \right)^2.
\end{equation}
The $(1-\alpha)$-level approximate CI is then formed by $\ybb \pm t_{k-1,\alpha/2} \sqrt{\varhat(\ybb)}$, where $t_{k-1,\alpha/2}$ denotes the $1-\alpha/2$ quantile of the Student's t distribution with $k-1$ degrees of freedom.  
Under appropriate conditions, $n\varhat(\ybb)$ is a consistent estimator of $\sigma^2$.

@@

The sample variance estimator is updated to
\begin{equation} \label{eq:overlap_variance_def}
	\vart{\ybb} = \frac{1}{(\tfrac{n}{m}-1)(n - m + 1)}\sum_{j=1}^{n-m+1} (\yb_j(m) - \ybb)^2.
\end{equation}
Note that the corresponding overlapping batch means are no longer independent. 
However, by simply reusing the data in this fashion---and seemingly in a contradictory way that induces dependence---one obtains a better variance estimator. 
\citet{Meketon1984} show that: (i) the overlapping variance estimator has nearly the same bias as the standard nonoverlapping variance estimator, and (ii) the overlapping estimator has only two-thirds of the asymptotic variance of the nonoverlapping one.  
That is, 
\begin{equation} \label{eq:asym_var}
	\frac{ \var{\vart{\ybb}} }{ \var{\varhat(\ybb)} } \rightarrow \frac{2}{3}
\end{equation}
in the limit as batch size $m$ and the number of batches $(n/m)$ tend to infinity.  
Note that the degrees of freedom in (\ref{eq:overlap_variance_def}) is slightly different than in \citep{Meketon1984}.  
The degrees of freedom in (\ref{eq:overlap_variance_def}) make $\vart{\ybb}$ an unbiased estimator for i.i.d.\ data for all finite $m$ and $n$ with the same asymptotic benefits \citep{Song1992}.  
The $(1-\alpha)$-level approximate CI is similarly formed by $\ybb \pm t_{3(k-1)/2,\alpha/2} \sqrt{\vart{\ybb}}$, with an adjustment to the degrees of freedom.

@@

Taking $n$ observations of the random variables $\xit$ and splitting them into $k$ nonoverlapping batches of size $m$, define 
\begin{align} 
	\gb_j  & = \frac{1}{m} \sum_{i = 1}^{m} f(\xh, \xit^{m(j-1)+i}) - \min_{\x \in X} \frac{1}{m} \sum_{i=1}^{m} f(\x, \xit^{m(j-1)+i}), \ \ \ \ j = 1,2, \dots, k \label{eq:mrp_gap}\\
         & = \frac{1}{m} \sum_{i = 1}^{m} \left[ f(\xh, \xit^{m(j-1)+i}) -  f(\xs_j, \xit^{m(j-1)+i}) \right], \ \ \ \ j = 1,2, \dots, k, \notag
\end{align}
where $\xs_j$ denotes an optimal solution to (\ref{eq:sto_prog_m}) of batch $j$ given on the right-hand side of (\ref{eq:mrp_gap}).    
Since we are assessing the quality of a given solution $\xh \in \X$, we suppress it from the notation $\gb_j$.
The optimality gap estimator $\gb_j$ is similar to $\yb_j$ with individual observations $y^{m(j-1)+i} = f(\xh,\xit^{m(j-1)+i}) - f(\xs_j,\xit^{m(j-1)+i})$, for $i=1,\ldots,m$ and $j=1,\ldots, k$.
Notice that, in this case, $y^i$ not only depends on observation $i$ of batch $j$, $\xit^{m(j-1)+i}$, but all the observations in that batch $\xit^{m(j-1)+1},\xit^{m(j-1)+2},\ldots,\xit^{m(j-1)+m}$ because $\xs_j$ is the optimal solution to (\ref{eq:sto_prog_m}) given in (\ref{eq:mrp_gap}).   
As before, after $k$ estimates are obtained, the overall mean of these estimates $\gbb = \frac{1}{k} \sum_{j=1}^k \gb_j$ provides a point estimator of the optimality gap.  
The sample variance is obtained as in (\ref{eq:var}) by $\varh{\gbb} = \frac{1}{k} \frac{1}{k-1} \sum_{j=1}^k (\gb_j - \gbb)^2$, which results in an approximate $(1-\alpha)$-level one-sided CI $\left[0, \gbb + t_{k-1,\alpha} \sqrt{\varh{\gbb}} \right]$.

Notice that the minimization in the second term on the right-hand side of (\ref{eq:mrp_gap}) gives rise to a biased gap estimator (recall the upper bound on the optimality gap).  
Thus, we can expect that the true probability of the optimality gap residing within the CI to be greater than the $1 - \alpha$ suggested by the above calculation.  
This is shown empirically in \citep{Bayraksan2006}, which also investigates additional methods for using a smaller number of replications (e.g., 1 or 2) with an alternative variance estimator to compute a CI.  
See also \citep{bayraksan_morton_09,partani2006jackknife,partani_07} for variations of MRP aimed to reduce bias and variance.

@@

As before, let $m$ denote the batch size, $n$ the total sample size and $k = \lfloor \tfrac{n}{m} \rfloor$ be the number of nonoverlapping batches.  
In this paper, we use the {\it batch nonoverlap parameter} $1 \leq \gamma \leq m$ to denote how much neighboring batches do {\it not} overlap.  
For instance, $\gamma = m$ corresponds to the case of nonoverlapping batches and $\gamma = 1$ corresponds to the maximally overlapping case of \citet{Meketon1984}.  
In general, $\gamma$ can take any integer value between and including $[1,m]$. 
Let $\lfloor \cdot \rfloor$ be the greatest integer that is less than or equal to its argument. 
The sample mean of each batch estimator is calculated similarly, 
$
\yb_j = \frac{1}{m} \sum_{i=1}^m y^{\gamma(j-1) + i},\ \  j = 1, 2, \dots, \nb, 
$
where $\nb$ is the number of batches used given $n$, $m$, and $\gamma$.  
The sample variance estimator given in (\ref{eq:overlap_variance_def}) is changed to
\begin{equation} \label{eq:variance_gamma}
	\vartg{\ybb} = \frac{1}{\left( \tfrac{n}{m} - 1 \right) \left( \nb \right)}  \sum_{j=1}^{\nb} (\yb_j - \ybb)^2.
\end{equation}
Note that with $\gamma=1$, (\ref{eq:variance_gamma}) reduces to (\ref{eq:overlap_variance_def}) and with $\gamma=m$ and $n=mk$, (\ref{eq:variance_gamma}) reduces to (\ref{eq:var}).

@@

The internal optimization in the batches in (\ref{eq:gbar}) makes a straightforward statistical analysis of the behavior of the estimators difficult.  
To overcome this problem, we introduce the following unbiased optimality gap estimators:
\begin{align}
	\db_{lj} & = \frac{1}{m_l} \sum_{i=1}^{m_l} \left[ f(\xh,\xit^{\gamma_l(j-1)+i}) - f(\xs,\xit^{\gamma_l(j-1)+i}) \right],\ \ \ j = 1, 2, \dots, \nbl, \label{eq:dbar} \\
	\dbb_l & = \frac{1}{n_l} \sum_{i=1}^{n_l} \left[ f(\xh,\xiti) - f(\xs,\xiti) \right], \label{eq:dbb} \\
	VD_l & = \frac{1}{\left( \tfrac{n_l}{m_l} - 1 \right) \left(\nbl\right)} \sum_{j=1}^{\nbl} (\db_{lj} - \dbb_l)^2. \label{eq:vd}
\end{align}
These are essentially the same as the variably overlapping batches estimators in \S \ref{ssec:overlap} with $y^i = f(\xh,\xiti) - f(\xs,\xiti)$.  
These are also defined identically to the original estimators (\ref{eq:gbar})--(\ref{eq:vg}) with the exception that $\xs_j$ from (\ref{eq:gbar}) and (\ref{eq:gbb}) is replaced by an optimal solution $\xs$ in (\ref{eq:dbar}) and (\ref{eq:dbb}).  
With $\xh$ and $\xs$ fixed, estimators (\ref{eq:dbar})--(\ref{eq:vd}) have the same statistical properties as variably overlapping batches estimators in \S \ref{ssec:overlap}.  
Note that the optimal solution $\xs$ is not known. 
However, the estimators (\ref{eq:dbar})--(\ref{eq:vd}) are used only to show convergence of (\ref{eq:gbar})--(\ref{eq:vg}); they are not necessary for carrying out the OMRP algorithm.  
Throughout the rest of this section, we assume that the optimality gap estimators (\ref{eq:gbar})--(\ref{eq:vg}) and their nonoptimized counterparts (\ref{eq:dbar})--(\ref{eq:vd}) use identical samples and values of $n_
l$, $m_l$, and $\gamma_l$.  \smallskip

@@

\begin{proof} When the sampling problems used in the calculation of $\gbb$ all have optimal solutions that solve (SP), i.e., when $\xs_{j} = \xs$, for all $j=1,2,\ldots, \nbl$, then $\gbb_l = \dbb_l$. 
Therefore, probability of the complement of this event provides an upper bound on the probability that $\dbb_l - \gbb_l \neq 0$. 
Let $E^C$ denote the complement of event $E$. 
Using this upper bound and continuing on, we obtain
	\begin{align}
		\p{\dbb_l - \gbb_l \neq 0} & \leq \p{\left( \bigcap_{j=1}^{\nbl} \{\xs_{j} = \xs\} \right)^C} \notag \\
		& = \p{ \bigcup_{j=1}^{\nbl} \{\xs_{j} \neq \xs\}} \notag \\
% 		& \leq \sum_{j=1}^{\nbl} \p{ \xs_j \neq \xs } \notag \\
		& \leq \left( \nbl\right) \p{ \xs_{m_l} \neq \xs }. \notag
	\end{align}
	With a batching structure in which $m_l = \lfloor n_l^{r} \rfloor$ for some $0<r<1$,  $\gamma_l \geq 1$, and by A3, the desired result follows.
\end{proof}

@@

\begin{proof}
	(i) By Lemma \ref{lem:gbb_prob}, $\p{\dbb_l - \gbb_l \neq 0}$ $\rightarrow 0$ exponentially fast.  
        Thus $\sum_l \p{\dbb_l - \gbb_l \neq 0} < \infty$.  
        The desired result follows by the Borel-Cantelli Lemma.
%
(ii) This proof is nearly identical to above.  
Note that
	\[
		\p{n_l VD_l -n_l VG_l \neq 0} \leq \p{\left( \bigcap_{j=1}^{\nbl} \{\xs_{j} = \xs\} \right)^C}
	\]
	and by the proof of Lemma \ref{lem:gbb_prob}, $\p{n_l VD_l -n_l VG_l \neq 0} \rightarrow 0$ exponentially fast, and thus $n_l VG_l$ is a consistent estimator of $\sigma^2_{\xh}$, if $n_l VD_l$ is, by the Borel-Cantelli Lemma.
\end{proof}

@@

\begin{proof}
	For simplicity of exposition, we drop the subscript $l$ and the $\frac{n_l}{m_l}$ for now and add them in later.  
        Further, we use abbreviations $n_b = \nbl$ to denote the number of batches and $dt = \left(\tfrac{n_l}{m_l} - 1\right) \left(\nbl\right)$ to denote the denominator term in $VG_l$ and $VD_l$. Note that
	\begin{align}
		\e{(nVG - nVD)^2} & = \left( \frac{n}{dt} \right)^2 \e{ \left( \sum_{j=1}^{n_b} \left[ (\gb_j - \gbb)^2 - (\db_j - \dbb)^2 \right] \right)^2 } \notag \\
		& \leq \left( \frac{n}{dt} \right)^2 \e{ n_b \sum_{j=1}^{n_b} \left[ (\gb_j - \gbb)^2 - (\db_j - \dbb)^2 \right]^2 } \label{eq:vvar_conv_pf_1}
	\end{align}
	by the Cauchy-Schwartz inequality, where the second vector in the Cauchy-Schwartz inequality is taken as a vector of ones.  
        By examining the batch structure in Figure \ref{fig:overlap_nonint} and the definition of $\gbb$ in (\ref{eq:gbb}), one can see that the terms of the expectation in (\ref{eq:vvar_conv_pf_1}) are not identically distributed.  
        We will separate these into identically distributed ``center'' batches and ``fringe'' batches.  
        For example, in Figure \ref{fig:overlap_nonint}, only batch 4 is a center batch, while the others are fringe batches.  
        Notice in Figure \ref{fig:overlap_nonint}, we see $\left\lceil \gammab^{-1} \right\rceil = 3$ sets of nonoverlapping batches $\{1,4,7\}$, $\{2,5\}$, and $\{3,6\}$, each of which has two fringe batches.  
        In general, there are $\left\lceil \gammab^{-1} \right\rceil$ sets of nonoverlapping batches, each of which will have two fringe batches, and $n_b - 2\left\lceil \gammab^{-1} \right\rceil$ center batches.

>>>>

We wish to assess the quality of a candidate solution $\xh \in X$, which may have been found in any way.
Assessing solution quality is important in practice because as (\ref{eq:sto_prog}) typically cannot be solved exactly, one only has an approximate solution $\xh$ without verification of its quality.
Assessing solution quality is also a critical component of stopping criteria in algorithms designed to (approximately) solve (\ref{eq:sto_prog}).
We define the quality of $\xh$ by its optimality gap, $\e{f(\xh,\xit)} - \zs$. 
The lower the optimality gap, the higher the quality of the solution; and a zero optimality gap implies that the solution is optimal. 
We assume that the candidate solution $\xh$ is given as input to our procedures, and hence fixed, throughout this paper.

@@

Our aim is to apply the idea of overlapping batches to MRP.  
Overlapping could help to obtain better variance estimators with simple data reuse, resulting in improved CIs.
%Especially for problems where sampling error is dominant, this simple data reuse method could yield better estimators.
We note several differences in this setting compared to the classical simulation setting.  
In simulation output analysis, the point of interest is that of estimating the variance of the sample mean of a covariance stationary process.  
We are interested in estimating the variance of an optimality gap estimator.  
An optimality gap estimator not only has a sample mean $(\frac{1}{m} \sum_{i=1}^m f(\xh,\xiti))$ but also a minimized sample mean $(\zs_m)$.  
Minimization changes the statistical properties of sample means, complicating the asymptotic analysis.  
%For instance, the central limit theorem may not hold for a minimized sample mean, even though it holds for each $\x \in \X$.  
%Therefore, analysis of the asymptotic benefits of overlapping becomes more complicated.
We overcome this technical difficulty by approximating the optimality gap estimators by their nonoptimized counterparts (see \S \ref{sec:theory}).  
The nonoptimized counterparts are (regular) overlapping batch means estimators (without any optimization present); therefore, they have the desired statistical properties of approximately the same bias and lower variances compared to their nonoverlapping versions. 
We establish convergence of the optimality gap estimators that use overlapping to their nonoptimized counterparts. 
Another difference between our setting and simulation output analysis is that once the data are generated through a simulation, they can be reused without much additional computational effort to obtain the OBM variance estimator.  
In this setting, due to the need to solve a sampling problem (SP$_m$), the computational effort may increase with data reuse.  
This can be alleviated in two ways. 
First, near-optimal variance reduction can be obtained by partially overlapping the batches \citep{Welch1987,Song1992}.  
Partial overlapping results in a smaller number of batches; hence, fewer optimization problems need to be solved.
Second, in many solution methods, warm-starting can be used to quickly solve the sampling approximations with overlapping samples, considerably reducing solution time.
In this paper, we consider variable amounts of overlap and examine this both theoretically and empirically. 
% THIS LAST SENTENCE NEEDS TO CHANGE!

@@

Nonoverlapping batch means takes $n$ observations, $y^1, \dots, y^n$ of the stochastic process and splits them into $k$ batches of size $m$, where $k = \frac{n}{m}$ (for simplicity, assume, for now, $n = mk$).  
The sample mean of each batch is then computed as
$
	\yb_j = \frac{1}{m} \sum_{i=1}^{m} y^{m(j-1)+i}, j = 1,2, \dots, k,
$
and the overall sample mean $\ybb = \frac{1}{k} \sum_{j=1}^k \yb_j = \frac{1}{n} \sum_{i=1}^n y^i$ provides a point estimator of $\mu$.  
The variance estimator of the batch means, which is $m/n$ times the sample variance, is calculated as
\begin{equation} \label{eq:var}
	\varhat(\ybb) = \frac{m}{n}\frac{1}{k-1} \sum_{j=1}^k \left( \yb_j - \ybb \right)^2.
\end{equation}
The $(1-\alpha)$-level approximate CI is then formed by $\ybb \pm t_{k-1,\alpha/2} \sqrt{\varhat(\ybb)}$, where $t_{k-1,\alpha/2}$ denotes the $1-\alpha/2$ quantile of the Student's t distribution with $k-1$ degrees of freedom.  
Under appropriate conditions, $n\varhat(\ybb)$ is a consistent estimator of $\sigma^2$.

@@

The variance estimator is updated to
\begin{equation} \label{eq:overlap_variance_def}
	\vart{\ybb} = \frac{1}{(\tfrac{n}{m}-1)(n - m + 1)}\sum_{j=1}^{n-m+1} (\yb_j(m) - \ybb)^2.
\end{equation}
The corresponding overlapping batch means are no longer independent; however, by simply reusing the data in this fashion---and seemingly in a contradictory way that induces dependence---one obtains a better variance estimator. 
\citet{Meketon1984} show that: (i) the overlapping variance estimator has nearly the same bias as the standard nonoverlapping variance estimator, and (ii) the overlapping estimator has only two-thirds of the asymptotic variance of the nonoverlapping one.  
That is, 
\begin{equation} \label{eq:asym_var}
	\frac{ \var{\vart{\ybb}} }{ \var{\varhat(\ybb)} } \rightarrow \frac{2}{3}
\end{equation}
in the limit as batch size $m$ and the number of batches $(n/m)$ tend to infinity.  
Note that the degrees of freedom in (\ref{eq:overlap_variance_def}) is slightly different than in \citep{Meketon1984}.  
The degrees of freedom in (\ref{eq:overlap_variance_def}) make $\vart{\ybb}$ an unbiased estimator for i.i.d.\ data for all finite $m$ and $n$ with the same asymptotic benefits \citep{Song1992}.  
The $(1-\alpha)$-level approximate CI is similarly formed by $\ybb \pm t_{3(k-1)/2,\alpha/2} \sqrt{\vart{\ybb}}$, with an adjustment to the degrees of freedom.

@@

Taking $n$ observations of the random variables $\xit$ and splitting them into $k$ nonoverlapping batches of size $m$, define 
\begin{align} 
	\gb_j  & = \frac{1}{m} \sum_{i = 1}^{m} f(\xh, \xit^{m(j-1)+i}) - \min_{\x \in X} \frac{1}{m} \sum_{i=1}^{m} f(\x, \xit^{m(j-1)+i}), \ \ \ \ j = 1,2, \dots, k \label{eq:mrp_gap}\\
         & = \frac{1}{m} \sum_{i = 1}^{m} \left[ f(\xh, \xit^{m(j-1)+i}) -  f(\xs_j, \xit^{m(j-1)+i}) \right], \ \ \ \ j = 1,2, \dots, k, \notag
\end{align}
where $\xs_j$ denotes an optimal solution to (\ref{eq:sto_prog_m}) of batch $j$ given on the right-hand side of (\ref{eq:mrp_gap}).    
Since we are assessing the quality of a given solution $\xh \in \X$, we suppress it from the notation $\gb_j$.
The optimality gap estimator $\gb_j$ is similar to $\yb_j$ with individual observations $y^{m(j-1)+i} = f(\xh,\xit^{m(j-1)+i}) - f(\xs_j,\xit^{m(j-1)+i})$, for $i=1,\ldots,m$ and $j=1,\ldots, k$.
Notice that in this case, $y^i$ not only depends on observation $i$ of batch $j$, $\xit^{m(j-1)+i}$, but all the observations in that batch $\xit^{m(j-1)+1},\xit^{m(j-1)+2},\ldots,\xit^{m(j-1)+m}$ because $\xs_j$ is the optimal solution to (\ref{eq:sto_prog_m}) given in (\ref{eq:mrp_gap}).   
As before, after $k$ estimates are obtained, the overall mean of these estimates $\gbb = \frac{1}{k} \sum_{j=1}^k \gb_j$ provides a point estimator of the optimality gap.  
The variance estimator is obtained as in (\ref{eq:var}) by $\varh{\gbb} = \frac{1}{k} \frac{1}{k-1} \sum_{j=1}^k (\gb_j - \gbb)^2$, which results in an approximate $(1-\alpha)$-level one-sided CI $\left[0, \gbb + t_{k-1,\alpha} \sqrt{\varh{\gbb}} \right]$.

The minimization in the second term on the right-hand side of (\ref{eq:mrp_gap}) gives rise to a biased gap estimator (recall the upper bound on the optimality gap).  
Thus, we can expect that the true probability of the optimality gap residing within the CI to be greater than the $1 - \alpha$ suggested by the above calculation.  
This is shown empirically in \citep{Bayraksan2006}, which also investigates additional methods for using a smaller number of replications (e.g., 1 or 2) with an alternative variance estimator to compute a CI.  
See also \citep{bayraksan_morton_09,partani2006jackknife,partani_07} for variations of MRP aimed to reduce bias and variance.

@@

As before, let $m$ denote the batch size, $n$ the total sample size and $k = \lfloor \tfrac{n}{m} \rfloor$ be the number of nonoverlapping batches.  
In this paper, we use the {\it batch nonoverlap parameter} $1 \leq \gamma \leq m$ to denote how much neighboring batches do {\it not} overlap.  
For instance, $\gamma = m$ corresponds to the case of nonoverlapping batches and $\gamma = 1$ corresponds to the maximally overlapping case of \citet{Meketon1984}.  
In general, $\gamma$ can take any integer value between and including $[1,m]$. 
Let $\lfloor \cdot \rfloor$ be the greatest integer that is less than or equal to its argument. 
The sample mean of each batch estimator is calculated similarly, 
$
\yb_j = \frac{1}{m} \sum_{i=1}^m y^{\gamma(j-1) + i},\ \  j = 1, 2, \dots, \nb, 
$
where $\nb$ is the number of batches used given $n$, $m$, and $\gamma$.  
The variance estimator given in (\ref{eq:overlap_variance_def}) is changed to
\begin{equation} \label{eq:variance_gamma}
	\vartg{\ybb} = \frac{1}{\left( \tfrac{n}{m} - 1 \right) \left( \nb \right)}  \sum_{j=1}^{\nb} (\yb_j - \ybb)^2.
\end{equation}
Note that with $\gamma=1$, (\ref{eq:variance_gamma}) reduces to (\ref{eq:overlap_variance_def}) and with $\gamma=m$ and $n=mk$, (\ref{eq:variance_gamma}) reduces to (\ref{eq:var}).

@@

The internal optimization in the batches in (\ref{eq:gbar}) makes a straightforward statistical analysis of the behavior of the estimators difficult.  
To overcome this problem, we introduce the following unbiased optimality gap estimators:
\begin{align}
	\db_{lj} & = \frac{1}{m_l} \sum_{i=1}^{m_l} \left[ f(\xh,\xit^{\gamma_l(j-1)+i}) - f(\xs,\xit^{\gamma_l(j-1)+i}) \right],\ \ \ j = 1, 2, \dots, \nbl, \label{eq:dbar} \\
	\dbb_l & = \frac{1}{n_l} \sum_{i=1}^{n_l} \left[ f(\xh,\xiti) - f(\xs,\xiti) \right], \label{eq:dbb} \\
	VD_l & = \frac{1}{\left( \tfrac{n_l}{m_l} - 1 \right) \left(\nbl\right)} \sum_{j=1}^{\nbl} (\db_{lj} - \dbb_l)^2. \label{eq:vd}
\end{align}
These are essentially the same as the variably overlapping batches estimators in \S \ref{ssec:overlap} with $y^i = f(\xh,\xiti) - f(\xs,\xiti)$.  
These are also defined identically to the original estimators (\ref{eq:gbar})--(\ref{eq:vg}) with the exception that $\xs_j$ from (\ref{eq:gbar}) and (\ref{eq:gbb}) is replaced by an optimal solution $\xs$ in (\ref{eq:dbar}) and (\ref{eq:dbb}).  
With $\xh$ and $\xs$ fixed, estimators (\ref{eq:dbar})--(\ref{eq:vd}) have the same statistical properties as variably overlapping batches estimators in \S \ref{ssec:overlap}.  
The optimal solution $\xs$ is not known; however, the estimators (\ref{eq:dbar})--(\ref{eq:vd}) are used only to show convergence of (\ref{eq:gbar})--(\ref{eq:vg}) and not necessary for carrying out the OMRP algorithm.  
Throughout the rest of this section, we assume that the optimality gap estimators (\ref{eq:gbar})--(\ref{eq:vg}) and their nonoptimized counterparts (\ref{eq:dbar})--(\ref{eq:vd}) use identical samples and values of $n_
l$, $m_l$, and $\gamma_l$.  \smallskip

@@

\begin{proof} When all the sampling problems used in the calculation of $\gbb$ have optimal solutions that solve (SP), i.e., when $\xs_{j} = \xs$, for all $j=1,2,\ldots, \nbl$, then $\gbb_l = \dbb_l$. 
Therefore, probability of the complement of this event provides an upper bound on the probability that $\dbb_l - \gbb_l \neq 0$. 
Let $E^C$ denote the complement of event $E$. 
Using this upper bound and continuing on, we obtain
	\begin{align}
		\p{\dbb_l - \gbb_l \neq 0} & \leq \p{\left( \bigcap_{j=1}^{\nbl} \{\xs_{j} = \xs\} \right)^C} \notag \\
		& = \p{ \bigcup_{j=1}^{\nbl} \{\xs_{j} \neq \xs\}} \notag \\
% 		& \leq \sum_{j=1}^{\nbl} \p{ \xs_j \neq \xs } \notag \\
		& \leq \left( \nbl\right) \p{ \xs_{m_l} \neq \xs }. \notag
	\end{align}
	With a batching structure in which $m_l = \lfloor n_l^{r} \rfloor$ for some $0<r<1$,  $\gamma_l \geq 1$, and by A3, the desired result follows.
\end{proof}

@@

\begin{proof}
	(i) By Lemma \ref{lem:gbb_prob}, $\p{\dbb_l - \gbb_l \neq 0}$ $\rightarrow 0$ exponentially fast.  
        Thus $\sum_l \p{\dbb_l - \gbb_l \neq 0} < \infty$.  
        The desired result follows by the Borel-Cantelli Lemma.
%
(ii) This proof is nearly identical to above. 
Similar to the proof of Lemma \ref{lem:gbb_prob}, we have
	\[
		\p{n_l VD_l -n_l VG_l \neq 0} \leq \p{\left( \bigcap_{j=1}^{\nbl} \{\xs_{j} = \xs\} \right)^C}.
	\]
	Following the same logic, $\p{n_l VD_l -n_l VG_l \neq 0} \rightarrow 0$ exponentially fast. 
Thus, $n_l VG_l$ is a consistent estimator of $\sigma^2_{\xh}$, if $n_l VD_l$ is, by the Borel-Cantelli Lemma.
\end{proof}

@@

\begin{proof}
	For simplicity of exposition, we drop the subscript $l$ and the $\frac{n_l}{m_l}$ for now and add them in later.  
        Further, we use abbreviations $n_b = \nbl$ to denote the number of batches and $dt = \left(\tfrac{n_l}{m_l} - 1\right) \left(\nbl\right)$ to denote the denominator term in $VG_l$ and $VD_l$. We have
	\begin{align}
		\e{(nVG - nVD)^2} & = \left( \frac{n}{dt} \right)^2 \e{ \left( \sum_{j=1}^{n_b} \left[ (\gb_j - \gbb)^2 - (\db_j - \dbb)^2 \right] \right)^2 } \notag \\
		& \leq \left( \frac{n}{dt} \right)^2 \e{ n_b \sum_{j=1}^{n_b} \left[ (\gb_j - \gbb)^2 - (\db_j - \dbb)^2 \right]^2 } \label{eq:vvar_conv_pf_1}
	\end{align}
	by the Cauchy-Schwartz inequality, where the second vector in the Cauchy-Schwartz inequality is taken as a vector of ones.  
        By examining the batch structure in Figure \ref{fig:overlap_nonint} and the definition of $\gbb$ in (\ref{eq:gbb}), one can see that the terms of the expectation in (\ref{eq:vvar_conv_pf_1}) are not identically distributed.  
        We will separate these into identically distributed ``center'' batches and ``fringe'' batches.  
        For example, in Figure \ref{fig:overlap_nonint}, only batch 4 is a center batch, while the others are fringe batches.  
        Notice in Figure \ref{fig:overlap_nonint}, we see $\left\lceil \gammab^{-1} \right\rceil = 3$ sets of nonoverlapping batches $\{1,4,7\}$, $\{2,5\}$, and $\{3,6\}$, each of which has two fringe batches.  
        In general, there are $\left\lceil \gammab^{-1} \right\rceil$ sets of nonoverlapping batches, each of which will have two fringe batches, and $n_b - 2\left\lceil \gammab^{-1} \right\rceil$ center batches.
